---
title: "04_DESeq2_macs"
output: html_document
date: "2023-11-11"
---

```{r}
library(tidyverse)
library(cowplot)
library(edgeR)
library(reshape2)
library(SingleCellExperiment)
library(pheatmap)
#library(apeglm)
library(png)
library(DESeq2)
library(RColorBrewer)
library(data.table)
library(FactoMineR)
library(factoextra)  
library(tidyverse)
library(pheatmap)
library(DESeq2)
library(RColorBrewer)
library(viridis)  
library(ggplot2)
library(grid)
library(ggrepel)
library(ggrepel)
library(Cairo)
```

#PCA
```{r}
# Transform counts for data visualization
rld <- vst(deseq, blind=TRUE)

# Plot PCA
p <- DESeq2::plotPCA(
  rld,
  ntop = 500,
  intgroup = c("group.id")
)
p
ggsave(
  filename = "./outs/LTA_vs_WT/Neu_pos_Epi/pca_by_group_id.pdf",
  plot = p,
  width = 6,
  height = 5
)

p <- DESeq2::plotPCA(
  rld,
  ntop = 500,
  intgroup = c("sample.id")
)
p
ggsave(
  filename = "./outs/LTA_vs_WT/Neu_pos_Epi/pca_by_sample_id.pdf",
  plot = p,
  width = 6,
  height = 5
)
```

```{r}
library(RColorBrewer)

#plot the PCA
pcaPlot <- plotPCA(rld, ntop = 500, intgroup = c("group.id","sample.id"), returnData = TRUE)
# Convert to ggplot object

# Set your desired figure size and font size
fig_width <- 6
fig_height <- 4
font_size <- 3 # Adjust as needed
point_size <- 3 # Adjust as needed

# Define colors for group.id (adjust the colors as needed)
# Get unique levels of group.id
group_levels <- unique(pcaPlot$group.id)

# Generate a color palette
num_colors <- length(group_levels)
colors <- sample(brewer.pal(max(3, min(9, num_colors)), "Set2"), num_colors)

# Create a named vector of colors
group_colors <- setNames(colors, group_levels)

ggplot(pcaPlot, aes(x = PC1, y = PC2, color = group.id, label = sample.id)) +
    geom_point(size = point_size) +
    geom_text_repel(size = font_size, 
                    box.padding = 0.8, 
                    point.padding = 1,
                    segment.size = 0.5) +
    scale_color_manual(values = group_colors) +
    theme(legend.position="none", 
          text = element_text(size = font_size),
          panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = font_size),
          axis.text.y = element_text(size = font_size),
          axis.title = element_text(size = font_size))


ggsave(filename = "pca_by_group_id_annotated.pdf", width = 8, height = 6, dpi = 300, path = './outs/LTA_vs_WT/Neu_pos_Epi/')
```

#Hierarchical clustering
```{r}
# Extract the rlog matrix from the object and compute pairwise correlation values
rld_mat <- assay(rld)
rld_cor <- cor(rld_mat)

# Plot heatmap
h <- pheatmap(rld_cor, annotation = cluster_metadata[, c("group.id", "sample.id"), drop=F])
h
cairo_pdf(file = './outs/LTA_vs_WT/Neu_pos_Epi/heatmap_clustering.pdf', width = 10, height = 6)
print(h)
dev.off()
```

#Running DESeq2
```{r}
# Run DESeq2 differential expression analysis
deseq <- DESeq(deseq)
# Plot dispersion estimates
#We can check the fit of the DESeq2 model to our data by looking at the plot of dispersion estimates.
#plotDispEsts(deseq)
#ggsave(filename = "DEGseq_plot.pdf", width = 12, height = 9, dpi = 300, path = '../four_bone_combined_analysis/20221021_old_vs_young_macs_outs//bulk_DEGs_outs/')
```

#Results
```{r}
#Let’s compare the stimulated group relative to the control:
# Output results of Wald test for contrast for pos vs neg
#method 1 design the exp and ctl groups -----------------------------------------------------------------------------------------
levels(cluster_metadata$group.id)[1]
levels(cluster_metadata$group.id)[2]
levels(cluster_metadata$group.id)[3]
levels(cluster_metadata$group.id)[4]

#the following contrast matrix can only compare two groups
contrast <- c("group.id", 
              levels(cluster_metadata$group.id)[2],
              levels(cluster_metadata$group.id)[1]
              )

res <- results(deseq, 
               contrast = contrast,
               alpha = 0.05)
res
```

#Table of results for all genes
```{r}
# Turn the results object into a tibble for use with tidyverse functions
res_tbl <- res %>%
        data.frame() %>%
        rownames_to_column(var="gene") %>%
        as_tibble()


# Check results output
res_tbl
res_tbl <- na.omit(res_tbl)
# Write all results to file
write.csv(res_tbl,
          quote = FALSE, 
          row.names = FALSE,
          './outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/res_tbl.csv')

saveRDS(res_tbl, './data/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/res_tbl.rds')
```

Table of results for significant genes
```{r}
# Set thresholds
pvalue_cutoff <- 0.05

# Subset the significant results
sig_res <- dplyr::filter(res_tbl, pvalue < pvalue_cutoff) %>%
        dplyr::arrange(pvalue)

# Check significant genes output
sig_res

write.csv(sig_res,
          quote = FALSE, 
          row.names = FALSE,
          './outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/sig_res.csv')

saveRDS(sig_res, './data/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/sig_res.rds')
```

```{r, normolize}
normalized_counts <- counts(deseq, normalized = TRUE)
```

#Scatterplot of normalized expression of top 20 most significant genes
##Now that we have identified the significant genes, we can plot a scatterplot of the top 20 significant genes. This plot is a good check to make sure that we are interpreting our fold change values correctly, as well.
```{r}
## ggplot of top genes
##fitering parameters
log2FC_cutoff = log2(1.5)
pvalue_cutoff = 0.05

##select the differential gene list 
need_DEG <- res_tbl[,c(1,3,6)] #select log2FoldChange, p value information
colnames(need_DEG) <- c('gene','log2FoldChange','p_value') 

need_DEG$significance  <- as.factor(ifelse(need_DEG$p_value < pvalue_cutoff & abs(need_DEG$log2FoldChange) > log2FC_cutoff, ifelse(need_DEG$log2FoldChange > log2FC_cutoff ,'UP','DOWN'),'NOT'))

title <- paste0(' Up :  ',nrow(need_DEG[need_DEG$significance =='UP',]) ,
                     '\n Down : ',nrow(need_DEG[need_DEG$significance =='DOWN',]),
                     '\n FoldChange >= ',round(2^log2FC_cutoff,3))

need_DEG$delabel <- NA
need_DEG$delabel[need_DEG$significance != "NOT"] <- row.names(need_DEG)[need_DEG$significance != "NOT"]

g <- ggplot(data=need_DEG, 
            aes(x=log2FoldChange, y=-log10(p_value), 
                color=significance, label=delabel)) +
  #dots and background
  geom_point(alpha=0.4, size=1) +
  theme_classic()+ #grid
  xlab("log2 ( FoldChange )") + 
  ylab("-log10 ( P.value )") +
  ggtitle( title ) +
  scale_colour_manual(values = c('blue','grey','red'))+ 
  scale_fill_brewer(palette="Set1")+
  geom_vline(xintercept = c(-log2FC_cutoff,log2FC_cutoff),lty=4,col="grey",lwd=0.5) +
  geom_hline(yintercept = -log10(pvalue_cutoff),lty=4,col="grey",lwd=0.5) +
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin=unit(c(2,2,2,2),'lines'), #top, right, bottom, left
        legend.title = element_blank(), 
        legend.position="right")
g

ggsave(g,filename = './outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/vocanol_no_labeling.pdf',width =8.5,height =8.5, dpi= 600)
```

```{r, with dot labeled}
library(dplyr)
library(ggplot2)
library(ggrepel)

# Separate UP and DOWN groups
up_genes <- need_DEG %>% filter(significance == "UP") %>% arrange(p_value) %>% head(100)
down_genes <- need_DEG %>% filter(significance == "DOWN") %>% arrange(p_value) %>% head(100)

# Combine back the top genes
top_genes <- rbind(up_genes, down_genes)

# Update the delabel column to only include labels for the top genes
need_DEG$delabel <- NA
need_DEG$delabel[need_DEG$gene %in% top_genes$gene] <- need_DEG$gene[need_DEG$gene %in% top_genes$gene]

# Plot
v <- ggplot(data=need_DEG, aes(x=log2FoldChange, y=-log10(p_value), col=significance, label=delabel)) +
    geom_point() + 
    theme_minimal() +
    geom_text_repel(aes(colour = significance)) +  # Use color aesthetics
    scale_color_manual(values=c("blue", "grey", "red")) +
    geom_vline(xintercept=c(-0.6, 0.6), lty=2, col="grey", lwd=0.5) +
    geom_point(alpha=0.4, size=1) +
    theme_classic() +
    xlab("log2 ( Fold Change )") + 
    ylab("-log10 ( P value )") +
    ggtitle(title) +
    scale_colour_manual(values = c('blue','grey','red'))+ 
    geom_vline(xintercept = c(-log2FC_cutoff, log2FC_cutoff), lty=2, col="grey", lwd=0.5) +
    geom_hline(yintercept = -log10(pvalue_cutoff), lty=2, col="grey", lwd=0.5) +
    theme(plot.title = element_text(hjust = 0.5), 
          plot.margin=unit(c(2,2,2,2),'lines'), 
          legend.title = element_blank(), 
          legend.position="right")
v
# Save the plot
ggsave(v, filename = 'volcano_with_labeling.pdf', width =10, height =10, dpi = 300, path = './outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/')
```

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)

# =========================
# User-defined parameters
# =========================

# Example:
# log2FC_cutoff <- log2(1.5)
# pvalue_cutoff <- 0.05

# Keywords for genes to highlight and label
# These can be partial gene names
highlight_keywords <- c(
  'Cdh2', 'Vim', 'Fap', 'Mmp2', 'Snai2', 'Twist1', 'Zeb2', 'Lox',
  'Serpine1', 'Wnt5a', 'Ctnnb1', 'Acta2', 'Foxc2', 'Snai1', 'Zeb1',
  'Mmp3', 'Mmp9', 'Fn1', 'Itgb6', 'Sox10', 'Cclc', 'Cdh1',
  'Epcam', 'Ocln', 'Cldn3', 'Cldn4', 'Cldn7', 'Krt8', 'Krt18',
  'Krt19', 'Dsp', 'Tjp1', 'Gata3', 'F11r'
)

highlight_color <- "#2f9765"

# TRUE = ignore upper/lower case
ignore_case_highlight <- FALSE

# TRUE = only label highlighted genes that are significant
# FALSE = label all highlighted genes
label_highlight_only_if_significant <- TRUE

# =========================
# Helper function: partial keyword matching
# =========================

match_keywords <- function(genes, keywords, ignore.case = FALSE) {
  keywords <- keywords[!is.na(keywords) & keywords != ""]
  
  if (length(keywords) == 0) {
    return(rep(FALSE, length(genes)))
  }
  
  sapply(
    genes,
    function(g) {
      any(
        vapply(
          keywords,
          function(k) grepl(k, g, fixed = TRUE, ignore.case = ignore.case),
          logical(1)
        )
      )
    }
  )
}

# =========================
# Prepare DEG table
# =========================

need_DEG <- as.data.frame(need_DEG)

need_DEG$p_value <- as.numeric(need_DEG$p_value)
need_DEG$log2FoldChange <- as.numeric(need_DEG$log2FoldChange)

# Avoid Inf when p_value is 0
need_DEG$p_plot <- pmax(need_DEG$p_value, .Machine$double.xmin)

# Determine significance
need_DEG$significance <- as.factor(
  ifelse(
    need_DEG$p_value < pvalue_cutoff &
      abs(need_DEG$log2FoldChange) > log2FC_cutoff,
    ifelse(
      need_DEG$log2FoldChange > log2FC_cutoff,
      "UP",
      "DOWN"
    ),
    "NOT"
  )
)

need_DEG$significance <- factor(
  need_DEG$significance,
  levels = c("DOWN", "NOT", "UP")
)

# =========================
# Identify keyword-highlighted genes
# =========================

need_DEG$highlight <- ifelse(
  match_keywords(
    genes = need_DEG$gene,
    keywords = highlight_keywords,
    ignore.case = ignore_case_highlight
  ),
  "highlight",
  "normal"
)

# Genes to label: only keyword-matched genes
if (label_highlight_only_if_significant) {
  label_df <- need_DEG %>%
    filter(highlight == "highlight", significance != "NOT")
} else {
  label_df <- need_DEG %>%
    filter(highlight == "highlight")
}

print("Highlighted genes:")
print(need_DEG$gene[need_DEG$highlight == "highlight"])

# =========================
# Plot title
# =========================

title <- paste0(
  " Up :  ", nrow(need_DEG[need_DEG$significance == "UP", ]),
  "\n Down : ", nrow(need_DEG[need_DEG$significance == "DOWN", ]),
  "\n FoldChange ≥ ", round(2^log2FC_cutoff, 3)
)

# =========================
# Volcano plot
# =========================

v <- ggplot(
  data = need_DEG,
  aes(
    x = log2FoldChange,
    y = -log10(p_plot),
    color = significance
  )
) +
  # All non-highlighted genes remain in the plot
  geom_point(
    data = subset(need_DEG, highlight == "normal"),
    alpha = 0.4,
    size = 1
  ) +
  
  # Highlighted genes shown in green
  geom_point(
    data = subset(need_DEG, highlight == "highlight"),
    aes(
      x = log2FoldChange,
      y = -log10(p_plot)
    ),
    color = highlight_color,
    size = 2.5,
    alpha = 1,
    inherit.aes = FALSE
  ) +
  
  # Only highlighted genes are annotated
  geom_text_repel(
    data = label_df,
    aes(
      x = log2FoldChange,
      y = -log10(p_plot),
      label = gene
    ),
    color = highlight_color,
    size = 4,
    max.overlaps = Inf,
    inherit.aes = FALSE
  ) +
  
  scale_color_manual(
    values = c(
      "DOWN" = "blue",
      "NOT" = "grey",
      "UP" = "red"
    )
  ) +
  
  geom_vline(
    xintercept = c(-log2FC_cutoff, log2FC_cutoff),
    lty = 2,
    col = "grey",
    lwd = 0.5
  ) +
  
  geom_hline(
    yintercept = -log10(pvalue_cutoff),
    lty = 2,
    col = "grey",
    lwd = 0.5
  ) +
  
  theme_classic() +
  xlab("log2 ( Fold Change )") +
  ylab("-log10 ( P value )") +
  ggtitle(title) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.margin = unit(c(2, 2, 2, 2), "lines"),
    legend.title = element_blank(),
    legend.position = "right"
  )

v

# =========================
# Save plot
# =========================

#ggsave(  plot = v,  filename = "volcano_with_keyword_highlighted_genes.pdf",  width = 10,  height = 10,  dpi = 300,  path = "./outs/LTA_vs_WT/cDC/")
```

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)

# =========================
# User-defined parameters
# =========================

# Example:
# log2FC_cutoff <- log2(1.5)
# pvalue_cutoff <- 0.05

# Exact gene names to highlight and label
highlight_genes <- c(
  'Cdh2', 'Vim', 'Fap', 'Mmp2', 'Snai2', 'Twist1', 'Zeb2', 'Lox',
  'Serpine1', 'Wnt5a', 'Ctnnb1', 'Acta2', 'Foxc2', 'Snai1', 'Zeb1',
  'Mmp3', 'Mmp9', 'Fn1', 'Itgb6', 'Sox10', 'Cclc', 'Cdh1',
  'Epcam', 'Ocln', 'Cldn3', 'Cldn4', 'Cldn7', 'Krt8', 'Krt18',
  'Krt19', 'Dsp', 'Tjp1', 'Gata3', 'F11r'
)

highlight_color <- "#2f9765"

# =========================
# Prepare DEG table
# =========================

need_DEG <- as.data.frame(need_DEG)

need_DEG$p_value <- as.numeric(need_DEG$p_value)
need_DEG$log2FoldChange <- as.numeric(need_DEG$log2FoldChange)

# Avoid Inf when p_value is 0
need_DEG$p_plot <- pmax(need_DEG$p_value, .Machine$double.xmin)

# Determine significance for background point colors
need_DEG$significance <- as.factor(
  ifelse(
    need_DEG$p_value < pvalue_cutoff &
      abs(need_DEG$log2FoldChange) > log2FC_cutoff,
    ifelse(
      need_DEG$log2FoldChange > log2FC_cutoff,
      "UP",
      "DOWN"
    ),
    "NOT"
  )
)

need_DEG$significance <- factor(
  need_DEG$significance,
  levels = c("DOWN", "NOT", "UP")
)

# =========================
# Exact gene-name matching
# =========================

need_DEG$highlight <- ifelse(
  need_DEG$gene %in% highlight_genes,
  "highlight",
  "normal"
)

# Label all exact-matched highlighted genes,
# regardless of significance threshold
label_df <- need_DEG %>%
  filter(highlight == "highlight")

print("Highlighted and annotated genes:")
print(label_df$gene)

print("Genes requested but not found in need_DEG:")
print(setdiff(highlight_genes, need_DEG$gene))

# =========================
# Plot title
# =========================

title <- paste0(
  " Up :  ", nrow(need_DEG[need_DEG$significance == "UP", ]),
  "\n Down : ", nrow(need_DEG[need_DEG$significance == "DOWN", ]),
  "\n FoldChange ≥ ", round(2^log2FC_cutoff, 3)
)

# =========================
# Volcano plot
# =========================

v <- ggplot(
  data = need_DEG,
  aes(
    x = log2FoldChange,
    y = -log10(p_plot),
    color = significance
  )
) +
  # All non-highlighted genes remain in the plot
  geom_point(
    data = subset(need_DEG, highlight == "normal"),
    alpha = 0.4,
    size = 1
  ) +
  
  # Highlighted exact-match genes in green
  geom_point(
    data = subset(need_DEG, highlight == "highlight"),
    aes(
      x = log2FoldChange,
      y = -log10(p_plot)
    ),
    color = highlight_color,
    size = 2.5,
    alpha = 1,
    inherit.aes = FALSE
  ) +
  
  # Annotate all highlighted genes regardless of threshold
  geom_text_repel(
    data = label_df,
    aes(
      x = log2FoldChange,
      y = -log10(p_plot),
      label = gene
    ),
    color = highlight_color,
    size = 4,
    max.overlaps = Inf,
    inherit.aes = FALSE
  ) +
  
  scale_color_manual(
    values = c(
      "DOWN" = "blue",
      "NOT" = "grey",
      "UP" = "red"
    )
  ) +
  
  geom_vline(
    xintercept = c(-log2FC_cutoff, log2FC_cutoff),
    lty = 2,
    col = "grey",
    lwd = 0.5
  ) +
  
  geom_hline(
    yintercept = -log10(pvalue_cutoff),
    lty = 2,
    col = "grey",
    lwd = 0.5
  ) +
  
  theme_classic() +
  xlab("log2 ( Fold Change )") +
  ylab("-log10 ( P value )") +
  ggtitle(title) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.margin = unit(c(2, 2, 2, 2), "lines"),
    legend.title = element_blank(),
    legend.position = "right"
  )

v

# =========================
# Save plot
# =========================

#ggsave(plot = v,filename = "volcano_with_EMT_gene_highlighting.pdf", height = 10,dpi = 300, path = "./outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/")
```


```{r, Heatmap}
# Heatmap
## Extract normalized counts for significant genes only
sig_counts <- normalized_counts[rownames(normalized_counts) %in% sig_res$gene, ]

## Set a color-blind friendly palette
heat_colors <- rev(brewer.pal(14, "PuOr"))

# Replace NA, NaN, and Inf with a specific value, such as 0
#sig_counts[is.na(sig_counts) | is.nan(sig_counts) | is.infinite(sig_counts)] <- 0

## Run pheatmap using the metadata data frame for the annotation
hp1 <- pheatmap(sig_counts, 
         color = heat_colors, 
         cluster_rows = TRUE, 
         show_rownames = FALSE,
         annotation = cluster_metadata[, c("sample.id", "group.id", "cell.id")], 
         border_color = NA, 
         fontsize = 7, 
         scale = "row", 
         fontsize_row = 10, 
         height = 20,
         angle_col = 45)

ggsave(hp1, filename = 'heatmap_sig_genes.pdf',width =8,height =16, dpi = 300, path = './outs/LTA_vs_WT/Neu_pos_Epi/B6J-wt_vs_B6J-LTA/')

```

```{r, selected genes for heatmap plot}
# Define genes
genes_of_interest <- c(  'Cdh2', 'Vim', 'Fap', 'Mmp2', 'Snai2', 'Twist1', 'Zeb2', 'Lox',
  'Serpine1', 'Wnt5a', 'Ctnnb1', 'Acta2', 'Foxc2', 'Snai1', 'Zeb1',
  'Mmp3', 'Mmp9', 'Fn1', 'Itgb6', 'Sox10', 'Cclc', 'Cdh1',
  'Epcam', 'Ocln', 'Cldn3', 'Cldn4', 'Cldn7', 'Krt8', 'Krt18',
  'Krt19', 'Dsp', 'Tjp1', 'Gata3', 'F11r')

# Subset result table
sig_res_subset <- res_tbl[res_tbl$gene %in% genes_of_interest, ]

# Keep gene order
sig_res_subset <- sig_res_subset[match(genes_of_interest, sig_res_subset$gene), , drop = FALSE]

# Extract normalized counts
sig_counts_subset <- normalized_counts[
  rownames(normalized_counts) %in% sig_res_subset$gene,
  ,
  drop = FALSE
]

# Reorder rows to match genes_of_interest
sig_counts_subset <- sig_counts_subset[
  genes_of_interest[genes_of_interest %in% rownames(sig_counts_subset)],
  ,
  drop = FALSE
]

library(pheatmap)
library(RColorBrewer)

heat_colors <- rev(brewer.pal(11, "PuOr"))

hp1 <- pheatmap(
  sig_counts_subset,
  color = heat_colors,
  cluster_rows = F,   # better for one gene
  show_rownames = TRUE,
  annotation_col = cluster_metadata[, c("sample.id", "group.id", "cell.id"), drop = FALSE],
  border_color = NA,
  fontsize = 7,
  scale = "row",
  fontsize_row = 10,
  angle_col = 45
)

hp1

```

```{r}
library(pheatmap)
library(RColorBrewer)

# Specify genes you want to label
genes_to_label <- c('Cdh2', 'Vim', 'Fap', 'Mmp2', 'Snai2', 'Twist1', 'Zeb2', 'Lox',
  'Serpine1', 'Wnt5a', 'Ctnnb1', 'Acta2', 'Foxc2', 'Snai1', 'Zeb1',
  'Mmp3', 'Mmp9', 'Fn1', 'Itgb6', 'Sox10', 'Cclc', 'Cdh1',
  'Epcam', 'Ocln', 'Cldn3', 'Cldn4', 'Cldn7', 'Krt8', 'Krt18',
  'Krt19', 'Dsp', 'Tjp1', 'Gata3', 'F11r')   # add more genes if needed

# Extract significant DEG counts
sig_counts <- normalized_counts[
  rownames(normalized_counts) %in% sig_res$gene,
  ,
  drop = FALSE
]

# Make darker / smoother PuOr color palette
heat_colors <- rev(colorRampPalette(brewer.pal(11, "PuOr"))(20))

# Define breaks
breaks <- seq(-1.5, 1.5, length.out = length(heat_colors) + 1)

# Create row labels:
# only label genes in genes_to_label; otherwise blank
row_labels <- ifelse(
  rownames(sig_counts) %in% genes_to_label,
  rownames(sig_counts),
  ""
)

# Optional: check which genes are actually found
genes_found <- genes_to_label[genes_to_label %in% rownames(sig_counts)]
genes_not_found <- setdiff(genes_to_label, genes_found)

genes_found
genes_not_found

hp1 <- pheatmap(
  sig_counts,
  color = heat_colors,
  breaks = breaks,
  cluster_rows = TRUE,
  show_rownames = TRUE,        # must be TRUE for labels_row to show
  labels_row = row_labels,     # only selected genes are labeled
  annotation_col = cluster_metadata[, c("sample.id", "group.id"), drop = FALSE],
  border_color = NA,
  fontsize = 7,
  scale = "row",
  fontsize_row = 6
)

hp1
```


```{r}
#make the plot color darker, adjust the color scale
# Function to make colors darker
sig_counts <- normalized_counts[rownames(normalized_counts) %in% sig_res$gene, ]

## Set a color-blind friendly palette
heat_colors <- rev(brewer.pal(20, "PuOr"))


# Define the breaks for the colors, ensuring to cover the range -4 to 4
breaks <- seq(-1.5, 1.5, length.out = length(heat_colors) + 1)

# Run pheatmap without the unsupported arguments
hp1 <- pheatmap(sig_counts, 
        color = heat_colors, 
        breaks = breaks,  # Use the defined breaks
        cluster_rows = TRUE, 
        show_rownames = FALSE,
        annotation = cluster_metadata[, c("sample.id", "group.id")], 
        border_color = NA, 
        fontsize = 7, 
        scale = "row", 
        fontsize_row = 5)
hp1
ggsave(hp1, filename = 'heatmap_sig_genes.pdf',device = cairo_pdf, width =10,height =12, dpi = 300, path = './outs/B_virus_1_2_3_7_11_B6_vs_FVB/')
```




